Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics

### 2022-03-03 Thesis/01-chap1.Rmd
Comment up to Section 1.4

## AI: Four Approaches

@@

Based on the answer to the above questions, four major approaches to pursue AI have been established. These approaches can be summarized into a two by two table as shown in Figure \@ref(fig:twodimai), where the row is "Human" vs. "Rational", and the column is "Behaviour" vs. "Thought". Positioning at the top right cell, the **rational agent approach** aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved. In contrast, the **"laws of thought" approach** focus on understanding the logic behind the rationality. Closely related to **cognitive science**, the **cognitive modelling approach** attempts to express theories of human cognition as computer program to mimic the thought process of human. Lastly, the **Turing test approach** is built upon the famous **Turing test** proposed by @turing_computing_1950. The test can be roughly described as, whether a human can distinguish another human from a computer with written communications only. To pass the test, several capabilities of computer are required. This includes **natural language processing** for communication with human, **knowledge representation** for encoding knowledge, **automated reasoning** for derivation of conclusions and **machine learning** for improving AI automatically through experience and data. Some researchers argued that written communication is insufficient to demonstrate intelligence, and some degree of physical simulation of a person is still necessary. One such example is the **total Turing test** proposed by @harnad_other_1991. It adds $3$ new requirements to the list, including **computer vision**, **speech recognition** and **robotics**, which are response for interactions with the physical world. Notably, all $7$ required capabilities have become major subfields of AI today. And their development has made AI one of the fastest-growing fields in the 21st century [@russell_artificial_2002].

@@

With the development of AI, mature AI technologies, such as facial recognition and web recommendation system, have profoundly affected the way modern society operates and citizen's daily life. This is largely as a consequence of the huge investment in AI industry by the financial market in recent years. Further, the increasingly cheap computing cost and the massive amount of accessible e-commerce data produced in the Internet age provide the possibility for applying data-intensive AI models, which enables AI performance to reach new heights in history [@jordan_machine_2015]. Some AI systems have already been remarkably better than human in certain areas, e.g., game playing. AlphaGo and AlphaZero developed by the Google DeepMind team surpass all human Go players [@silver_general_2018].

@@

Predictive models are primarily evaluated by predictive accuracy with the use of validation and test data, but in predictive model diagnostics, especially model testing and tuning, data plots play an irreplaceable role. In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed by researchers, due to its intuitiveness and the possibility to provide unexpected discoveries which may be abstract and unquantifiable.

However, unlike confirmatory data analysis built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective. Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data. One such example is finding a separation between gene groups in a two-dimensional projection from a linear discriminant analysis where there is no difference in the expression levels between the gene groups [@roy_chowdhury_using_2015].

@@

Visual inference was first introduced by @buja_statistical_2009 as an inferential framework to extend confirmatory statistics to visual discoveries. This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing. Figure \@ref(fig:parallelism) outlines the parallelism between conventional tests and visual discovery.

@@

>>>>

## Artificial intelligence: Four Approaches

@@

Based on the answer to the above questions, four major approaches to pursue AI have been established. These approaches can be summarized into a two by two table as shown in Figure \@ref(fig:twodimai), where the row is "Human" vs. "Rational", and the column is "Behaviour" vs. "Thought". Positioning at the top right cell, the **rational agent approach** aims to build agent that perform mathematically perfect acts such that the best expected outcome can always be achieved. In contrast, the **"laws of thought" approach** focus on understanding the logic behind the rationality. Closely related to **cognitive science**, the **cognitive modelling approach** attempts to express theories of human cognition as computer program to mimic the thought process of human. Lastly, the **Turing test approach** is built upon the famous **Turing test** proposed by @turing_computing_1950. The test can be roughly described as, whether a human can distinguish another human from a computer with written communications only. To pass the test, several capabilities of computer are required. This includes **natural language processing** for communication with human, **knowledge representation** for encoding knowledge, **automated reasoning** for derivation of conclusions and **machine learning** for improving AI automatically through experience and data. Some researchers argued that written communication is insufficient to demonstrate intelligence, and some degree of physical simulation of a person is still necessary. One such example is the **total Turing test** proposed by @harnad_other_1991. It adds three new requirements to the list, including **computer vision**, **speech recognition** and **robotics**, which are response for interactions with the physical world. Notably, all seven required capabilities have become major subfields of AI today. And their development has made AI one of the fastest-growing fields in the 21st century [@russell_artificial_2002].

@@

With the development of AI, mature AI technologies, such as facial recognition and web recommendation system, have profoundly affected the way modern society operates and citizen's daily life. This is largely as a consequence of the huge investment in AI industry by the financial market in recent years. Further, the increasingly cheap computing cost and the massive amount of accessible e-commerce data produced in the Internet age provide the possibility for applying data-intensive AI models, which enables AI performance to reach new heights in history [@jordan_machine_2015]. Some AI systems have already been remarkably better than humans in certain areas, e.g., game playing. AlphaGo and AlphaZero developed by the Google DeepMind team surpass professional human Go players [@silver_general_2018].

@@

Predictive models are primarily evaluated by predictive accuracy with the use of validation and test data, but in predictive model diagnostics, especially model testing and tuning, data plots play an irreplaceable role. In these diagnostics, though numeric summaries are mostly available and some are even endorsed by finite or asymptotic properties, graphical representation of data is still preferred, or at least needed by researchers, due to its intuitiveness and the possibility for unexpected discoveries which may be abstract and unquantifiable.

However, unlike confirmatory data analysis built upon rigorous statistical procedures, e.g., hypothesis testing, visual diagnostics relies on graphical perception - human’s ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984], which is to some extent subjective. Further, visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data. One such example is finding [ET: an over-interpretation of the] separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurence [@roy_chowdhury_using_2015].

@@

Visual inference was first introduced by @buja_statistical_2009 as an inferential framework to extend confirmatory statistics to visual discoveries. [ET: this idea was actually presented first in 1999 JSM talk "Inference for Data Visualization” by Buja et al. http://stat.wharton.upenn.edu/~buja/PAPERS/visual-inference.pdf -- as cited by Gelman in 2003 https://paperpile.com/shared/u0Gy7K] This framework redefines the test statistics, tests, null distribution, significance levels and $p$-value for visual discovery modelled on the confirmatory statistical testing. Figure \@ref(fig:parallelism) outlines the parallelism between conventional tests and visual discovery.

@@

[ET: Caption of Figure 1.3 requires fixing. E.g. what is null hypothesis referring to? What visual discovery are you referring to?]
